---
title: "Methodology - BrCa Screening"
author: "Ernesto Ulloa"
date: "September 7, 2015"
output: html_document
---

The first  4 sections of this markdown file define:

- Auxiliary functions to run the analysis 
- Population and detection parameters
- Effectiveness measured in DALYS
- Costs of each program
 

In the final section we run the cost-effectiveness analysis for all posible screening programs. 

# Auxiliary functions

First we load all necessary libraries and  auxiliary functions described in the *Auxiliary Functions* markdown file. 

```{r,warning=FALSE,message=FALSE}
library(RCurl)
library(foreign)
library(xtable)
library(knitr)
library(nlme)
library(KMsurv)
library(ggplot2)
library(plyr)
library(reshape2)

url <-'https://rawgit.com/ernesto-up/Reproducibility-Cost-Effectiveness/master/Auxiliary%20Functions/auxiliary-functions.R'
temp <- getURL(url,followlocation = TRUE, cainfo = system.file("CurlSSL", "cacert.pem", package = "RCurl"))
source(url)

```

# Population structure

In this section we load and define population data and parameters. 

## Data from National Population Counsil in Mexico

We load the data from the National Population Counsil and define the age groups.

```{r}
url <- "https://rawgit.com/ernesto-up/Reproducibility-Cost-Effectiveness/master/Data%20Sets/Conapo-Projections.csv"
temp <- getURL(url)                
conapo <- read.csv(textConnection(temp))
```

### Age Groups

Define age groups

```{r}
age.group.data.frame <- c('[25,40)','[40,45)','[45,50)','[50,55)','[55,60)','[60,65)','[65,70)','[70,75)')
age.group <- c(25,40,45,50,55,60,65,70,75)
```

### Mean of group age

Define the mean of each group age:

```{r}
mean.age.group <- data.frame(age.group = age.group.data.frame,mean.age.group = c(32.5,42.5,47.5,52.5,57.5,62.5,67.5,72.5))
```

## BrCa Incidence

We define the incidence rates by group age according to GLOBOCAN's estimates of the incidence rates.

```{r}
incidence <- data.frame(age.group = age.group.data.frame,incidence = c(11.7,58.7,84.3,106.5,118.6,119.4,116.6,108.8))
```

## Estimation of BrCa letality rates by cancer stage

We load the data from the patient cohort (n = 615) treated at the National Cancer Institute in Mexico covered financially by the Catastrophic Protection Fund with follow up from 2007 to 2013 

```{r,warning=FALSE,message=FALSE}
githubURL <-'https://rawgit.com/ernesto-up/Reproducibility-Cost-Effectiveness/master/Data%20Sets/Catastrophic-Protection-Fund-Data.RData'
load(url(githubURL))
```

### Transformation of variables

We convert the time to years and separate them by clinical stage:

```{r,warning=FALSE,message=FALSE}
protection.fund.data$obs.time.years <- as.numeric((protection.fund.data$tiempo.vida)/365.25)
data.by.stage <- split(protection.fund.data,protection.fund.data$etapa.clinica.2)
```

### Life tables by clinical stage

We calculate the life tables by stages: 

```{r,message=FALSE,warning=FALSE}
life.tables <- lapply(data.by.stage,function(x){
                            obs.time.years.redo <-floor(x$obs.time.years)
                            data.table <- data.frame(obs.time.years.redo,x$censura.vida)
                            die<-gsummary(data.table,sum,groups=obs.time.years.redo)
                            total<-gsummary(data.table, length, groups=obs.time.years.redo)
                            ltab.data<-cbind(die[,1:2], total[,2])
                            rm(obs.time.years.redo)
                            attach(ltab.data)
                            lt=length(obs.time.years.redo)
                            obs.time.years.redo[lt+1]=NA
                            nevent = x.censura.vida
                            nlost = total[,2] - x.censura.vida
                            n = nrow(x)
                            mytable <- lifetab(obs.time.years.redo, n, nlost, nevent)
                            detach(ltab.data)
                            mytable[,1:5]
                            })
```

#### Early Stage 

Resulting life table for patients in early BrCa stage

```{r,echo=FALSE}
kable(life.tables[[1]][1:6,])
```

#### Locally Advanced Stage 

Resulting life table for patients in locally advanced BrCa stage

```{r,echo=FALSE}
kable(life.tables[[2]][1:6,])
```

### Metastatic Stage 

Resulting life table for patients for mestastasic BrCa stage

```{r,echo=FALSE}
kable(life.tables[[3]][1:6,])
```

With the previous life tables we define the following data frame:

```{r,echo=FALSE}
survival <- data.frame(clinical.stage = c(rep('Early',5),rep('Locally Advanced',5),rep('Metastasic',5)),
rate.survival  = c(0.9940,0.9819,0.9634,0.9447,0.9384,0.9727,0.9116,0.8389,0.7835,0.7377,
                           0.6854,0.6020,0.4854,0.3712,0.2579),year = c(1:5,1:5,1:5))
  
survival$rate.survival <- round(survival$rate.survival,2)

survival <- ddply(survival,'clinical.stage',mutate,
                    acum.survival = c(rate.survival[1],
                             prod(rate.survival[1:2]),                           
                             prod(rate.survival[1:3]),
                             prod(rate.survival[1:4]), 
                             prod(rate.survival[1:5])))

survival$acum.survival <- round(survival$acum.survival,2)

survival <- ddply(survival,'clinical.stage',mutate,
                   acum.letality = c(1-acum.survival[1],
                                      acum.survival[1]- acum.survival[2],
                                      acum.survival[2] - acum.survival[3],
                                      acum.survival[3] - acum.survival[4],
                                      acum.survival[4] - acum.survival[5]))

survival$acum.letality <- round(survival$acum.letality,2)
```

## BrCa Null Scenario Distribution

We define the distribution of BrCa cases of patients who arrived at the National Institute of Cancer in Mexico during 2007 without screening. This is the BrCa distribution under the null scenario. 

```{r}
distribution <- data.frame(clinical.stage = c('Early','Locally Advanced','Metastasic'),distribution = c(.344,.521,.135))
```

# Breast Cancer Screening: Detection Parameters

## Mean Sojourn Time
Mean sojourn time (Duffy & Gabe, a partir de los 40) is defined for group ages of 40-49, 50-59, 60-69, 70-74 so we extrapolate for the 
25-40 group age

```{r,echo=FALSE}
x <- c(45,55,65,72)
y <- c(2.4,3.7,4.2,4)
lin.mst <- lm(y ~ x) 
qplot(x = x,y = y) + geom_point() + geom_smooth(method = 'lm',se = F,fullrange = T) + 
  scale_x_continuous(limits= c(20,73)) + xlab('Age in years') + ylab('Latent period in years')
pred <- data.frame(x = c(32.5))
predict(lin.mst, newdata = pred)
```

Define mean sojourn time with extrapolated results:

```{r}
mean.sojourn.time <- data.frame(mean.sojourn.time = c(1.96,2.4,2.4,3.7,3.7,4.2,4.2,4.0),
                                age.group  = age.group.data.frame)
```

## Sensitivity of the mammogram 

Sensitivity defined in (Otha et al.) is  defined  for the 40-49, 50-59, 60-69, 70-79 group ages, we use a linear model on the log of the sensitivity to extrapolate for the 25-40 age group.

```{r,echo=FALSE}
x <- c(45,55,65,75) 
y <- c(.698,.667,.773,.838)
lin.sens <- lm(log(y) ~ x)
z <- x
a <- log(y)
qplot(x = z,y = a) + geom_point() + geom_smooth(method = 'lm',formula = a ~ x ,se = F,fullrange = T) + 
  scale_x_continuous(limits= c(20,80)) + xlab('age') + ylab('sensitivity')
pred <- data.frame(x = c(32.5))
exp(predict(lin.sens, newdata = pred))
rm(list = c('pred','lin.mst','lin.sens','x','y','z'))
```

Define sensitivity with extrapolated results:

```{r,echo=FALSE}
sensitivity <- data.frame(age.group = age.group.data.frame,sensitivity= c(.611,.698,.698,.667,.667,.773,.773,.838))
```

# DALYS

We define the disability weights according to WHO and the Disability Weights Project Study.  

```{r,echo=FALSE}
disability.weights <- data.frame(stage = c('Early','Locally Advanced','Metastasic','Survival','Death'),
                                 disability.weights.first.year = c(.27,.37,.61,.05,1),
                                 disability.weights.second.year = c(.18,.30,.61,.05,1))
```

## Estimation of DALYS 

In this section we estimate the number of DALYS lost per each stage of BrCa. 

First, we define life expectacy according to the [National Institute of Statistics and Geography in Mexico](http://cuentame.inegi.org.mx/poblacion/esperanza.aspx?tema=P)

```{r}
life.expectancy <- 77.5
```

### DAYLS Early Stage:

```{r}
survival.early <- subset(survival,clinical.stage == 'Early')

YLD.early <- ddply(mean.age.group,'age.group',summarize,
                  YLD.early = 1*burden(N = 1,DW = subset(disability.weights,stage == 'Early')$disability.weights.first.year
                               ,A = mean.age.group,L = 1,a = mean.age.group) + 
                    survival.early$acum.survival[1]*burden(1,
                           subset(disability.weights,stage == 'Early')$disability.weights.second.year,
                           A = mean.age.group+1,L=1,a = mean.age.group) +
                    survival.early$acum.survival[2]*burden(1,
                           subset(disability.weights,stage == 'Early')$disability.weights.second.year,
                           A = mean.age.group+2,L=1,a = mean.age.group) +
                    survival.early$acum.survival[3]*burden(1,
                           subset(disability.weights,stage == 'Early')$disability.weights.second.year,
                           A = mean.age.group+3,L=1,a = mean.age.group) +
                    survival.early$acum.survival[4]*burden(1,
                           subset(disability.weights,stage == 'Early')$disability.weights.second.year,
                           A = mean.age.group+4,L=1,a = mean.age.group) +
                    survival.early$acum.survival[5]*burden(1,
                           subset(disability.weights,stage == 'Survival')$disability.weights.second.year,
                           A = mean.age.group+5,L=life.expectancy-(mean.age.group+5),a = mean.age.group)) 

YLL.early <- ddply(mean.age.group,'age.group',summarize,
                      YLL.early = 0*burden(N = 1,DW =1,A = mean.age.group,L = life.expectancy - mean.age.group,a = mean.age.group) + 
                        survival.early$acum.letality[1]*burden(1,DW= 1,
                               A = mean.age.group+1,L=life.expectancy-(mean.age.group+1),a = mean.age.group) +
                        survival.early$acum.letality[2]*burden(1,DW= 1,
                               A = mean.age.group+2,L=life.expectancy-(mean.age.group+2),a = mean.age.group) +
                        survival.early$acum.letality[3]*burden(1,DW= 1,
                               A = mean.age.group+3,L=life.expectancy-(mean.age.group+3),a = mean.age.group) +
                        survival.early$acum.letality[4]*burden(1,DW= 1,
                               A = mean.age.group+4,L=life.expectancy-(mean.age.group+4),a = mean.age.group) +
                        survival.early$acum.letality[5]*burden(1,DW= 1,
                               A = mean.age.group+5,L=life.expectancy-(mean.age.group+5),a = mean.age.group)) 
```

### DAYLS Locally Advanced: 

```{r}
survival.locally.advanced <- subset(survival,clinical.stage == 'Locally Advanced')

YLD.locally.advanced <- ddply(mean.age.group,'age.group',summarize,
                      YLD.locally.advanced = burden(N = 1,DW = 
                        subset(disability.weights,stage == 'Locally Advanced')$disability.weights.first.year,
                                   A = mean.age.group,L = 1,a = mean.age.group) + 
                        survival.locally.advanced$acum.survival[1]*burden(1,
                               subset(disability.weights,stage == 'Locally Advanced')$disability.weights.second.year,
                               A = mean.age.group+1,L=1,a = mean.age.group) +
                        survival.locally.advanced$acum.survival[2]*burden(1,
                               subset(disability.weights,stage == 'Locally Advanced')$disability.weights.second.year,
                               A = mean.age.group+2,L=1,a = mean.age.group) +
                        survival.locally.advanced$acum.survival[3]*burden(1,
                               subset(disability.weights,stage == 'Locally Advanced')$disability.weights.second.year,
                               A = mean.age.group+3,L=1,a = mean.age.group) +
                        survival.locally.advanced$acum.survival[4]*burden(1,
                               subset(disability.weights,stage == 'Locally Advanced')$disability.weights.second.year,
                               A = mean.age.group+4,L=1,a = mean.age.group) +
                        survival.locally.advanced$acum.survival[5]*burden(1,
                               subset(disability.weights,stage == 'Survival')$disability.weights.second.year,
                               A = mean.age.group+5,L=life.expectancy-(mean.age.group+5),a = mean.age.group)) 

YLL.locally.advanced <- ddply(mean.age.group,'age.group',summarize,
                      YLL.locally.advanced = 0*burden(N = 1,DW =1,A = mean.age.group,
                        L = life.expectancy- mean.age.group,a = mean.age.group) + 
                        survival.locally.advanced$acum.letality[1]*burden(1,DW= 1,
                          A = mean.age.group+1,L=life.expectancy-(mean.age.group+1),a = mean.age.group) +
                        survival.locally.advanced$acum.letality[2]*burden(1,DW= 1,
                          A = mean.age.group+2,L=life.expectancy-(mean.age.group+2),a = mean.age.group) +
                        survival.locally.advanced$acum.letality[3]*burden(1,DW= 1,
                          A = mean.age.group+3,L=life.expectancy-(mean.age.group+3),a = mean.age.group) +
                        survival.locally.advanced$acum.letality[4]*burden(1,DW= 1,
                          A = mean.age.group+4,L=life.expectancy-(mean.age.group+4),a = mean.age.group) +
                        survival.locally.advanced$acum.letality[5]*burden(1,DW= 1,
                          A = mean.age.group+5,L=life.expectancy-(mean.age.group+5),a = mean.age.group)) 
```

### Dalys Metastasic: 

```{r}
survival.metastasic <- subset(survival,clinical.stage == 'Metastasic')

YLD.metastasic<- ddply(mean.age.group,'age.group',summarize,
                      YLD.metastasic = burden(N = 1,DW = subset(disability.weights,stage == 'Metastasic')$disability.weights.first.year,
                                          A = mean.age.group,L = 1,a = mean.age.group) + 
                                            survival.metastasic$acum.survival[1]*burden(1,
                                              subset(disability.weights,stage == 'Metastasic')$disability.weights.second.year,
                                          A = mean.age.group+1,L=1,a = mean.age.group) +
                                            survival.metastasic$acum.survival[2]*burden(1,
                                              subset(disability.weights,stage == 'Metastasic')$disability.weights.second.year,
                                          A = mean.age.group+2,L=1,a = mean.age.group) +
                                            survival.metastasic$acum.survival[3]*burden(1,
                                              subset(disability.weights,stage == 'Metastasic')$disability.weights.second.year,
                                          A = mean.age.group+3,L=1,a = mean.age.group) +
                                            survival.metastasic$acum.survival[4]*burden(1,
                                              subset(disability.weights,stage == 'Metastasic')$disability.weights.second.year,
                                          A = mean.age.group+4,L=1,a = mean.age.group) +
                                            survival.metastasic$acum.survival[5]*burden(1,
                                              subset(disability.weights,stage == 'Survival')$disability.weights.second.year,
                                          A = mean.age.group+5,L=life.expectancy-(mean.age.group+5),a = mean.age.group)) 

YLL.metastasic <- ddply(mean.age.group,'age.group',summarize,
                                 YLL.metastasic = 0*burden(N = 1,DW =1,A = mean.age.group,
                                    L = life.expectancy- mean.age.group,a = mean.age.group) +
                                      survival.metastasic$acum.letality[1]*burden(1,DW= 1,
                                    A = mean.age.group+1,L=life.expectancy-(mean.age.group+1),a = mean.age.group) +
                                      survival.metastasic$acum.letality[2]*burden(1,DW= 1,
                                    A = mean.age.group+2,L=life.expectancy-(mean.age.group+2),a = mean.age.group) +
                                      survival.metastasic$acum.letality[3]*burden(1,DW= 1,
                                    A = mean.age.group+3,L=life.expectancy-(mean.age.group+3),a = mean.age.group) +
                                      survival.metastasic$acum.letality[4]*burden(1,DW= 1,
                                    A = mean.age.group+4,L=life.expectancy-(mean.age.group+4),a = mean.age.group) +
                                     survival.metastasic$acum.letality[5]*burden(1,DW= 1,
                                    A = mean.age.group+5,L=life.expectancy-(mean.age.group+5),a = mean.age.group)) 


total.DALY.early <- data.frame(age.group = YLD.early$age.group, 
                                  total.DALY.early  = YLD.early$YLD.early + YLL.early$YLL.early)
                            
total.DALY.advanced <- data.frame(age.group = YLD.locally.advanced$age.group, 
                            total.DALY.advanced = YLD.locally.advanced$YLD.locally.advanced + 
                            YLL.locally.advanced$YLL.locally.advanced)
                            
total.DALY.metastasic <-  data.frame(age.group = YLD.metastasic$age.group, 
                                total.DALY.metastasic  = YLD.metastasic$YLD.metastasic+ 
                                YLL.metastasic$YLL.metastasic)
                                

DALY <- merge(total.DALY.early,total.DALY.advanced)
DALY <- merge(DALY,total.DALY.metastasic)
```

# Costs of Screening and Treatment

Deine the cost of one mammography:

```{r,echo=FALSE}
cost.mammography.2015 <- 1186
```

Define costs per clinical stage considering infalation rate:

```{r,echo=FALSE}
costs.stage.2012 <- data.frame(costs = c(161736,225064,241923), 
                           clinical.stage = c('Early','Locally Advanced','Metastasic'))

rate.cost <- .03 

costs.stage.2015 <- data.frame(costs = costs.stage.2012$costs*(1 + rate.cost)^3,
                                clinical.stage = costs.stage.2012$clinical.stage)

costs <- data.frame(cost.mammography = 0, cost.stage.Early = 0, 
                     cost.stage.locally.advanced =0, cost.stage.metastasic=0,
                     period = seq(2016,2026,1))

costs$cost.mammography <- cost.mammography.2015*(1 + rate.cost)^(-1*seq(1,11,1))
costs$cost.stage.Early <- costs.stage.2015[1,1]*(1 + rate.cost)^(-1*seq(1,11,1))
costs$cost.stage.locally.advanced <- costs.stage.2015[2,1]*(1 + rate.cost)^(-1*seq(1,11,1))
costs$cost.stage.metastasic <- costs.stage.2015[3,1]*(1 + rate.cost)^(-1*seq(1,11,1))

```

## Mexican GDP
We define the mexican GDP according to the [world bank](http://data.worldbank.org/indicator/NY.GDP.PCAP.CD)

```{r,echo=FALSE}
GDP.Mexico <- 10307 
```

# Cost Effectiveness Analysis 

## Calculamos casos nuevos con tasas de GLOBOCAN y base de CONAPO:

```{r}
conapo.incidence <- population.incidence(incidence,age.group,conapo,mean.age.group)
conapo.incidence <- merge(conapo.incidence,DALY)
conapo.incidence <- merge(conapo.incidence,costs,id = 'period')
```

# Cost Effectiveness Analysis

In this last section we run the simulations to obtain the cost effectiveness estimates for all the screening programs. First we load the data.

```{r}

library(plyr)
library(reshape2)
library(ggplot2)

```

Define all the screening percentages for each age group: 

```{r}

porc.25.40 <- seq(0,2,1)*.01
porc.40.45 <- seq(0,15,5)*.01
porc.45.50 <- seq(0,18,6)*.01
porc.50.55 <- seq(0,20,5)*.01
porc.55.60 <- seq(0,25,5)*.01
porc.60.65 <- seq(0,20,5)*.01
porc.65.70 <- seq(0,18,6)*.01
porc.70.75 <- seq(0,2,2)*.01

```

Define periodicities of the programs:

```{r}
periodicity<- c(1,2,3)
```

Starting and end year:

```{r}
start<- 2016
end <- 2026
```

We create a data frame with all the information for each intervention and run the simulations. We store all the results in the same data frame. This last part of the markdown file is not evaluated. 

It takes approximately 40 minutes to run with:

MacBook Pro 
2.9 GHz Intel Core i5 
8 GB 1867 MHz DDR3
R version 3.2.0 (2015-04-16)

```{r,eval=FALSE}
 results.simulation <- data.frame(expand.grid(list = list(periodicity = periodicity,porc.25.40 = porc.25.40, porc.40.45 = porc.40.45,
                        porc.45.50 = porc.45.50, porc.50.55 = porc.50.55, porc.55.60 = porc.55.60,
                        porc.60.65 = porc.60.65, porc.65.70 = porc.65.70, porc.70.75 = porc.70.75)), total.dalys = 0,
                        total.cost.masto = 0, total.treatment = 0, total.cost = 0)

for(i in 1:dim(results.simulation)[1])
{
  percentage <- data.frame(percentage = as.numeric(results.simulation[i,2:9]),age.group=age.group.data.frame)
  periodicity <- as.numeric(results.simulation[i,'periodicity'])
  results <- resume.ce(percentage,periodicity,incidence,conapo.incidence,mean.sojourn.time,sensitivity,
                        start,end,age.group.data.frame)
  results.simulation[i,10:13] <- results
}

```







