---
title: "Results"
author: "Ernesto Ulloa"
date: "September 28, 2015"
output: html_document
---

# Load Data

First we load the results from the simulation study and the auxiliary functions:

```{r}
library(plyr)
library(reshape2)
library(ggplot2)
githubURL <-'https://rawgit.com/ernesto-up/Reproducibility-Cost-Effectiveness/master/Data%20Sets/Results%20Simulation.RData'
load(url(githubURL))

githubURL <- 'https://rawgit.com/ernesto-up/Reproducibility-Cost-Effectiveness/master/Data%20Sets/Simulation%20Parameters.RData'
load(url(githubURL))
```


# Null Model 

Define the null model when the simulation has no screening program

```{r}
null.model <- results.simulation[1,]
```


# Current scenario

Using data from the ENSANUT survey based on the question **During 2011 did you attend a health service unit to...** we define the current screening program:


```{r}
ensanut<- data.frame(periodicity = 2,porc.25.40 = .019,porc.40.45 = .096,porc.45.50 = .145,
                     porc.50.55 = .196, porc.55.60 = .189,porc.60.65 = .179, porc.65.70 = .145, 
                     porc.70.75 = .073,total.dalys = 0,
                     total.cost.masto = 0,total.treatment = 0,total.cost = 0)

percentage <- data.frame(percentage= as.numeric(ensanut[2:9]),age.group=age.group.data.frame)
periodicity <- as.numeric(ensanut[1,'periodicity'])
start<- 2016
end <- 2026
```

In order to compare the current program, we evaluate its cost effectiveness:

```{r}
results.ensanut <- resume.ce(percentage,periodicity,incidence,conapo.incidence,mean.sojourn.time,sensitivity,
                              start,end,age.group.data.frame)
```

# Costs and DALYS

Calculate the differences in costs and DALYS, between the null model and the rest of the models:

```{r}

results.ensanut$dif.cost <- results.ensanut$total.cost - null.model$total.cost
results.ensanut$dif.dalys <- null.model$total.dalys - results.ensanut$total.dalys 
results.ensanut$dif.cost.dollars.year <- (results.ensanut$dif.cost)/(11*15.22)
results.ensanut$dif.dalys.year <- results.ensanut$dif.dalys/10

results.simulation.2 <- results.simulation
results.simulation.2$dif.cost <- results.simulation$total.cost - null.model$total.cost
results.simulation.2$dif.dalys <- null.model$total.dalys - results.simulation.2$total.dalys 
results.simulation.2$dif.cost.dollars.year <- (results.simulation.2$dif.cost)/(11*15.22)
results.simulation.2$dif.dalys.year <- results.simulation.2$dif.dalys/10
results.simulation.2$Periodicity <- as.factor(with(results.simulation.2,ifelse(periodicity ==1,'Annual',
                                                                        ifelse(periodicity == 2,'Biennial','Triennial'))))
```
######################################################################################

# Encontramos las intervenciones que tengan menor cociente de cost efectividad

Find programs with the smallest cost-effectiveness ratio:


```{r}
results.simulation.3<- results.simulation.2
results.simulation.3$cost.efect<- results.simulation.2$dif.cost.dollars.year/results.simulation.2$dif.dalys.year
optimo <- results.simulation.3[which(results.simulation.3$cost.efect == min(results.simulation.3$cost.efect,na.rm = TRUE)),]
optimo$cost.efect.opt <- optimo$cost.efect
```


Find the expansion path:

```{r,warning=FALSE}
exit <- 0
optimo.temp <- optimo

while(exit == 0){
  
  temp <- results.simulation.3
  
  # remove interventions no longer in the expansion path
  temp <- temp[c(which(temp$dif.cost.dollars.year >= optimo.temp$dif.cost.dollars.year &
                         temp$dif.dalys.year >= optimo.temp$dif.dalys.year)),]
  
  if(dim(temp)[1] != 0){
  # encontramos nueva intervención óptima
  delta.y <- temp$dif.cost.dollars.year - optimo.temp$dif.cost.dollars.year
  delta.x  <- temp$dif.dalys.year - optimo.temp$dif.dalys.year
  temp$cost.efect.opt <- delta.y/delta.x
  print(min(temp$cost.efect.opt,na.rm = TRUE))
  optimo.temp <- temp[which(temp$cost.efect.opt == min(temp$cost.efect.opt,na.rm = TRUE)),]
  optimo <- rbind(optimo,optimo.temp)
  
  }else {
    exit <-  1
  } 
}

```

# Plot Results

We plot the results of the simulation:

```{r}
optimo <- optimo[which(optimo$Periodicity != 'Anual'),]
feasible <- optimo[12,]
objective <- optimo[13,]

cbPalette <- c("#E69F00", "#56B4E9", "#CC79A7","#009E73","#F0E442", "#0072B2","#D55E00")

theme_set(theme_bw(20))

g <- ggplot(results.simulation.2,aes(x = dif.dalys.year,y = dif.cost.dollars.year/10^6,colour = Periodicity))+ 
  ggtitle('Cost-Efficiency of Mammography Screening \n 2016-2026') + geom_point(alpha = .25) +
  theme(plot.title = element_text(size=20, face="bold", vjust=2))+
  geom_abline(intercept  = 0, slope = 3*GDP.Mexico/10^6,size =1.1,color = 'grey10',linetype = 'longdash') +  
  geom_abline(intercept  = 0, slope = GDP.Mexico/10^6,size = 1.1, color = 'grey10',linetype = 'longdash') + 
  xlab('DALYS Averted per year') + ylab('Annual Cost Millions of US$') + 
  coord_cartesian(xlim = c(0,50000))+ coord_cartesian(ylim = c(0,300)) + scale_color_manual(values = cbPalette) + 
  guides(colour = guide_legend(override.aes = list(alpha = 1))) + 
  geom_point(aes(x = results.ensanut$dif.dalys.year, y = results.ensanut$dif.cost.dollars.year/10^6), 
             colour = 'firebrick3' ,size = 3,shape = 23,fill = 'firebrick3')+
  annotate('text', x = results.ensanut$dif.dalys.year + 1200, y = results.ensanut$dif.cost.dollars.year/10^6-5, 
           label = 'Current Program', color = 'firebrick3',size = 6)+
  geom_line(data = optimo,aes(x = dif.dalys.year,y = dif.cost.dollars.year/10^6),colour = 'grey22')+
  geom_point(data = optimo,aes(x = dif.dalys.year,y = dif.cost.dollars.year/10^6),
             size = 3,shape = 23,colour = 'grey22', fill = 'grey22')+
  annotate('text', x = 7800, y = 22, 
           label = 'Expansion Path', color = 'grey22',size = 6)+
  geom_point(aes(x = feasible$dif.dalys.year, y = feasible$dif.cost.dollars.year/10^6),
             colour = 'orangered', size= 3,shape = 23, fill = 'orangered') +
  annotate('text', x = feasible$dif.dalys.year + 1000, y =  feasible$dif.cost.dollars.year/10^6,
           label = 'Feasible', color = 'orangered',size = 6) +
  geom_point(aes(x = objective$dif.dalys.year, y = objective$dif.cost.dollars.year/10^6),
             colour = 'forestgreen', size= 3,shape = 23, fill = 'forestgreen') + 
  annotate('text', x = objective$dif.dalys.year + 1350, y =  objective$dif.cost.dollars.year/10^6,
           label = 'Objetive', color = 'forestgreen',size = 6) + 
  annotate('text', x = 11000, y =  280,label = '3 veces GDP', color = 'grey22',size = 6) +
  annotate('text', x = 15000, y =  180,label = 'GDP', color = 'grey22',size = 6) 
  
```

```{r,echo=FALSE}
g
```

We compare the current intervention with the feasible intervention:

```{r}
feasible$dif.dalys.year - results.ensanut$dif.dalys.year
(feasible$dif.cost.dollars.year- results.ensanut$dif.cost.dollars.year)/10^6
```
